---
title: "Week 5 Discussion Section"
author: "Rachel Swick"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Learning Objectives
- Use `terra` functions `aggregate()` and `resample()` to create a new raster
- Use `terra` functions `as.polygons()` to convert a raster to a vector of polygons

## Get Started
```{r}
library(tidyverse)
library(sf)
library(terra)
library(spData)
library(spDataLarge)
```

```{r}
# Load raster data representing grain sizes with the three classes clay, silt and sand
grain <- terra::rast(system.file("raster/grain.tif", package = "spData"))
```

## Your Task

1. Subset points higher than 3100 meters in `nz_height` (from `spData`) and create a template raster with `rast()`, where the resolution is 3 km x 3 km for the extent of the subset dataset
- Count numbers of the highest points in each grid cell
- Find the maximum elevation in each grid cell

```{r}
# Create a subset of New Zealand elevation points over 3100 meters
nz_height_subset <- nz_height %>% 
  dplyr::filter(elevation > 3100)

# Create template: define the extent, resolution, and CRS based on nz_height_subset
nz_template <- rast(terra::ext(nz_height_subset), 
                    resolution = 3000, 
                    crs = terra::crs(nz_height_subset))
```

```{r}
# Count numbers of the highest points in each grid cell
nz_raster <- rasterize(nz_height_subset, 
                       nz_template, 
                       field = "elevation", 
                       fun = "length")

# Find the maximum elevation in each grid cell
plot(nz_raster, 
     main = "Number of Elevation Points > 3100 in Each Grid Cell")
plot(st_geometry(nz_height_subset), add = TRUE)
```

2. With the previous raster, complete the following:
- Aggregate the raster that counts the highest points in New Zealand/Aotearoa
- Reduce its geographic resolution by half, such that cells are 6 x 6 km
- Plot the result
- Resample back to the original resolution of 3 km x 3 km

```{r}
nz_raster_lower <- aggregate(nz_raster, 
                             fact = 2, 
                             fun = sum, 
                             na.rm = TRUE)

nz_raster_resample <- resample(nz_raster_lower,
                               nz_raster)

plots <- c(nz_raster, nz_raster_resample)
labs <- c("Original 6 x 6 km", "Resample 6 x 6 km")
plot(plots, main = labs)
```



3. Polygonize grain and filter to only keep squares that represent clay














